---
title: "Model Performance"
author: "Group"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
# Default knitting options
knitr::opts_chunk$set(echo=FALSE, # Whether to echo the code
                      # tidy=TRUE, # Nicely tidy up code
                      warning=FALSE, # No warnings please 
                      message=FALSE) # No messages please

options(warn=-1) 

# Suppress start up warnings when loading libraries
library <- function(...) {
  suppressPackageStartupMessages(base::library(...))
}
```

```{r libraries}
# Load in all libraries
library(tidyverse)
library(here) # directory referencing
library(stargazer) # used by John for nice display of model output
library(knitr) # used for nice display of tables
library(dplyr)
library(ConfusionTableR)
library(caret)
library(tidyr)
library(cvms)
```

```{r load data and models}
load(file = here("data", "model_data.Rdata"))
load(file = "models.Rdata")
```

# Part 1 of 3: LDL
## 1.1 GLM on LDL_status
```{r}
predicted_glm_LDL <- predict(glm_LDL, model_data, type = "response")
predicted_glm_LDL <- as.factor(round(predicted_glm_LDL, 0) + 1)
caret::confusionMatrix(predicted_glm_LDL, model_data$LDL_status)



png(here("plots", "glm_LDL_confusion_matrix.png"), width = 960, height = 960, pointsize = 24)
ConfusionTableR::binary_visualiseR(train_labels = predicted_glm_LDL,
                                   truth_labels= model_data$LDL_status,
                                   class_label1 = "Normal", 
                                   class_label2 = "Abnormal",
                                   quadrant_col1 = "#28ACB4", 
                                   quadrant_col2 = "#4397D2", 
                                   custom_title = "LDL Logistic Confusion Matrix", 
                                   text_col= "black")



```

## 1.2 Ordinal on LDL_band
```{r}
predicted_tbl_LDL_o <- 
  predict(ordinal_LDL, model_data, type = "p") %>% 
  as.tibble() %>% 
  mutate(id = row_number()) %>% 
  pivot_longer(cols = 1:8) %>% 
  group_by(id) %>% 
  mutate(max_value = max(value),
         is_predicted = max_value == value) %>% 
  filter(is_predicted) %>% 
  mutate(LDL_status_predicted = if_else(name > 5, 2, 1))

LDL_band_predicted_o <- 
  as.factor(as.vector(predicted_tbl_LDL_o$name)) %>% 
  # add all levels in case some are missing
  forcats::fct_expand("1", "2", "3", "4", "5", "6", "7", "8")

LDL_status_predicted_o <- 
  as.factor(as.vector(predicted_tbl_LDL_o$LDL_status_predicted)) %>% 
  forcats::fct_expand("1", "2")

caret::confusionMatrix(LDL_status_predicted_o, model_data$LDL_status)
caret::confusionMatrix(LDL_band_predicted_o, model_data$LDL_band)

png(here("plots", "ordinal_LDL_confusion_matrix.png"), width = 960, height = 960, pointsize = 24)
ConfusionTableR::binary_visualiseR(train_labels = LDL_status_predicted_o,
                                   truth_labels= model_data$LDL_status,
                                   class_label1 = "Normal", 
                                   class_label2 = "Abnormal",
                                   quadrant_col1 = "#28ACB4", 
                                   quadrant_col2 = "#4397D2", 
                                   custom_title = "LDL Ordinal Confusion Matrix", 
                                   text_col= "black")


```

### 1.2.2 ordinal LDL - nicely presented confusion matrix

```{r}

conf_mat <- cvms::confusion_matrix(targets = model_data$LDL_band,
                             predictions = LDL_band_predicted_o)
conf_mat$`Confusion Matrix`[[1]]

plot_confusion_matrix(
  conf_mat$`Confusion Matrix`[[1]],
  # font_counts = font(
  #   size = 10
  # ),
  class_order = as.character(c(8,7,6,5,4,3,2,1)),
  add_normalized = FALSE,
  add_col_percentages = FALSE,
  add_row_percentages = FALSE
)

caret_conf_mat <- caret::confusionMatrix(LDL_band_predicted_o, model_data$LDL_band)


caret_conf_mat$overall %>% 
  janitor::clean_names() %>% 
  as.list() %>% 
  as_tibble() %>% 
  dplyr::select(Accuracy = accuracy, "Highest Class Proportion" = accuracy_null, "Accuracy p Value" = accuracy_p_value) %>% 
  kable(digits = 2, format="pipe")
```


## 1.3 multinomial on LDL_band
```{r}
predicted_tbl_LDL_m <- 
  predict(multinom_LDL, model_data, type = "p") %>% 
  as.tibble() %>% 
  mutate(id = row_number()) %>% 
  pivot_longer(cols = 1:8) %>% 
  group_by(id) %>% 
  mutate(max_value = max(value),
         is_predicted = max_value == value) %>% 
  filter(is_predicted) %>% 
  mutate(LDL_status_predicted = if_else(name > 5, 2, 1))

LDL_band_predicted_m <- 
  as.factor(as.vector(predicted_tbl_LDL_m$name)) %>% 
  # add all levels in case some are missing
  forcats::fct_expand("1", "2", "3", "4", "5", "6", "7", "8")

LDL_status_predicted_m <- 
  as.factor(as.vector(predicted_tbl_LDL_m$LDL_status_predicted)) %>% 
  forcats::fct_expand("1", "2")

caret::confusionMatrix(LDL_status_predicted_m, model_data$LDL_status)



```


# Part 2 of 3: HDL
## 2.1 GLM on HDL_status
```{r}
predicted_glm_HDL <- predict(glm_HDL, model_data, type = "response")
predicted_glm_HDL <- as.factor(predicted_glm_HDL > 0.5)
caret::confusionMatrix(predicted_glm_HDL, model_data$HDL_status)
```


## 2.2 Ordinal on HDL_band
```{r}
predicted_tbl_HDL_o <- 
  predict(ordinal_HDL, model_data, type = "p") %>% 
  as.tibble() %>% 
  mutate(id = row_number()) %>% 
  pivot_longer(cols = 1:6) %>% 
  group_by(id) %>% 
  mutate(max_value = max(value),
         is_predicted = max_value == value) %>% 
  filter(is_predicted) %>% 
  mutate(HDLNTR_predicted = if_else(name > 5, 2, 1))

HDLRESB_predicted_o <- 
  as.factor(as.vector(predicted_tbl_HDL_o$name)) %>% 
  # add all levels in case some are missing
  forcats::fct_expand("1", "2", "3", "4", "5", "6")

HDLNTR_predicted_o <- 
  as.factor(as.vector(2 == predicted_tbl_HDL_o$HDLNTR_predicted)) 

caret::confusionMatrix(HDLNTR_predicted_o, model_data$HDL_status)
caret::confusionMatrix(HDLRESB_predicted_o, model_data$HDL_band)
```
### 2.2.2 ordinal HDL - nicely presented confusion matrix

```{r}

conf_mat <- cvms::confusion_matrix(targets = model_data$HDL_band,
                             predictions = HDLRESB_predicted_o)
conf_mat$`Confusion Matrix`[[1]]

plot_confusion_matrix(
  conf_mat$`Confusion Matrix`[[1]],
  # font_counts = font(
  #   size = 10
  # ),
  class_order = as.character(c(6,5,4,3,2,1)),
  add_normalized = FALSE,
  add_col_percentages = FALSE,
  add_row_percentages = FALSE
)

caret_conf_mat <- caret::confusionMatrix(HDLRESB_predicted_o, model_data$HDL_band)


caret_conf_mat$overall %>% 
  janitor::clean_names() %>% 
  as.list() %>% 
  as_tibble() %>% 
  dplyr::select(Accuracy = accuracy, "Highest Class Proportion" = accuracy_null, "Accuracy p Value" = accuracy_p_value) %>% 
  kable(digits = 3, format="pipe")
```

## 2.3 multinomial on HDL_band
```{r}
predicted_tbl_HDL_m <- 
  predict(multinom_HDL, model_data, type = "p") %>% 
  as.tibble() %>% 
  mutate(id = row_number()) %>% 
  pivot_longer(cols = 1:6) %>% 
  group_by(id) %>% 
  mutate(max_value = max(value),
         is_predicted = max_value == value) %>% 
  filter(is_predicted) %>% 
  mutate(HDLNTR_predicted = if_else(name > 5, 2, 1))

HDLRESB_predicted_m <- 
  as.factor(as.vector(predicted_tbl_HDL_m$name)) %>% 
  # add all levels in case some are missing
  forcats::fct_expand("1", "2", "3", "4", "5", "6")

HDLNTR_predicted_m <- 
  as.factor(as.vector(2 == predicted_tbl_HDL_m$HDLNTR_predicted)) 

caret::confusionMatrix(HDLNTR_predicted_m, model_data$HDL_status)
caret::confusionMatrix(HDLRESB_predicted_m, model_data$HDL_band)

```

### 2.3.2 multinomial HDL - nicely presented confusion matrix

```{r}

conf_mat <- cvms::confusion_matrix(targets = model_data$HDL_band,
                             predictions = HDLRESB_predicted_m)

plot_confusion_matrix(
  conf_mat$`Confusion Matrix`[[1]],
  # font_counts = font(
  #   size = 10
  # ),
  class_order = as.character(c(6,5,4,3,2,1)),
  add_normalized = FALSE,
  add_col_percentages = FALSE,
  add_row_percentages = FALSE
)

caret_conf_mat <- caret::confusionMatrix(HDLRESB_predicted_m, model_data$HDL_band)


caret_conf_mat$overall %>% 
  janitor::clean_names() %>% 
  as.list() %>% 
  as_tibble() %>% 
  dplyr::select(Accuracy = accuracy, "Highest Class Proportion" = accuracy_null, "Accuracy p Value" = accuracy_p_value) %>% 
  kable(digits = 3, format="pipe")
```


# Part 3 of 3: CHOL (Total cholesterol)
## 3.1 GLM on total_cholesterol_status
```{r}
predicted_glm_CHOL <- predict(glm_CHOL, model_data, type = "response")
predicted_glm_CHOL <- as.factor(round(predicted_glm_CHOL, 0) + 1)
caret::confusionMatrix(predicted_glm_CHOL, model_data$total_cholesterol_status)
```


## 3.2 Ordinal on total_cholesterol_band
```{r}

predicted_tbl_CHOL_o <- 
  predict(ordinal_CHOL, model_data, type = "p") %>% 
  as.tibble() %>% 
  mutate(id = row_number()) %>% 
  pivot_longer(cols = 1:8) %>% 
  group_by(id) %>% 
  mutate(max_value = max(value),
         is_predicted = max_value == value) %>% 
  filter(is_predicted) %>% 
  mutate(total_cholesterol_status_predicted = if_else(name > 5, 2, 1))

total_cholesterol_band_predicted_o <- 
  as.factor(as.vector(predicted_tbl_CHOL_o$name)) %>% 
  # add all levels in case some are missing
  forcats::fct_expand("1", "2", "3", "4", "5", "6", "7", "8")

total_cholesterol_status_predicted_o <- 
  as.factor(as.vector(predicted_tbl_CHOL_o$total_cholesterol_status_predicted)) %>% 
  forcats::fct_expand("1", "2")

caret::confusionMatrix(total_cholesterol_status_predicted_o, model_data$total_cholesterol_status)
caret::confusionMatrix(total_cholesterol_band_predicted_o, model_data$total_cholesterol_band)
```


## 3.3 multinomial on total_cholesterol_band
```{r}
predicted_tbl_CHOL_m <- 
  predict(multinom_CHOL, model_data, type = "p") %>% 
  as.tibble() %>% 
  mutate(id = row_number()) %>% 
  pivot_longer(cols = 1:8) %>% 
  group_by(id) %>% 
  mutate(max_value = max(value),
         is_predicted = max_value == value) %>% 
  filter(is_predicted) %>% 
  mutate(total_cholesterol_status_predicted = if_else(name > 5, 2, 1))

total_cholesterol_band_predicted_m <- 
  as.factor(as.vector(predicted_tbl_CHOL_m$name)) %>% 
  # add all levels in case some are missing
  forcats::fct_expand("1", "2", "3", "4", "5", "6", "7", "8")

total_cholesterol_status_predicted_m <- 
  as.factor(as.vector(predicted_tbl_CHOL_m$total_cholesterol_status_predicted)) %>% 
  forcats::fct_expand("1", "2")

caret::confusionMatrix(total_cholesterol_status_predicted_m, model_data$total_cholesterol_status)
caret::confusionMatrix(total_cholesterol_band_predicted_m, model_data$total_cholesterol_band)
```

