---
title: "Group 7 Supplementary R Analysis"
author: "Group 7"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Default knitting options
knitr::opts_chunk$set(echo=TRUE, # Whether to echo the code
                      tidy=TRUE, # Nicely tidy up code
                      warning=FALSE, # No warnings please 
                      message=FALSE) # No messages please

options(warn=-1) 
```

```{r libraries}
# Suppress start up warnings when loading libraries
library <- function(...) {
  suppressPackageStartupMessages(base::library(...))
}
# Load in all libraries
library(tidyverse)
library(here) # directory referencing
library(MASS) # ordinal regression
library(nnet) # multinomial regression
library(caret)
library(knitr) # used for nice display of tables
```

# Data transformations

```{r creating model data set}
# This contains tech_biom and tech_nutr from Assignment 1
load(file = here("data", "tech_data_selected.Rdata"))

# The final columns chosen for our analysis
required_columns <- c(
            'LDLNTR',
            'LDLRESB',
            'WHOLGR1N',
            'REFGRA1N',
            'SYSTOL',
            'SLPTIME',
            'BMISC',
            'AGEC',
            'EIBMR1',
            'PHDCMWBC',
            'EXLWTBC', 
            'HDLCHREB',
            'CHOLNTR',
            'CHOLRESB',
            'FIBRPER1',
            'FIBRPER2')

# Data transformations and making variables human-readable 
model_data <-  
  tech_nutr %>%
  inner_join(tech_biom, by = "ABSPID") %>% 
  dplyr::select(all_of(required_columns)) %>% 
  filter(AGEC >= 18,
         # dropping observations below Goldberg cut-off
         EIBMR1 >= 0.9) %>% 
  mutate(LDLNTR = droplevels(LDLNTR),
         LDLRESB = droplevels(LDLRESB),
         # abnormal HDL when HDLCHREB = 1 (< 1 mmol)
         HDLCHREB_BINOM = as.factor(as.numeric(HDLCHREB) != 1),
         HDLCHREB = droplevels(HDLCHREB),
         CHOLNTR = droplevels(CHOLNTR),
         CHOLRESB = droplevels(CHOLRESB),
         EXLWTBC = as.numeric(EXLWTBC),
         SYSTOL = as.numeric(SYSTOL),
         SLPTIME = as.numeric(SLPTIME),
         AGEC = as.numeric(AGEC),
         # avg energy from fibre across the two days
         FIBRPER = (FIBRPER1 + FIBRPER2)/2
         ) %>% 
  # removing the per-day fibre variables
  dplyr::select(-FIBRPER1, -FIBRPER2) %>%
  # dropping observations with NA for any variable, leaves us with 2405 rows
  drop_na() %>% 
  # renaming to human-readable names
  rename(LDL_status =	LDLNTR,
          LDL_band =	LDLRESB,
          WG =	WHOLGR1N,
          RG =	REFGRA1N,
          systolic_blood_pressure =	SYSTOL,
          sleep_time =	SLPTIME,
          BMI =	BMISC,
          age =	AGEC,
          metabolic_rate =	EIBMR1,
          weight =	PHDCMWBC,
          exercise_duration =	EXLWTBC,
          HDL_status =	HDLCHREB_BINOM,
          HDL_band =	HDLCHREB,
          total_cholesterol_status =	CHOLNTR,
          total_cholesterol_band =	CHOLRESB,
          fibre =	FIBRPER)
```


# LDL
## GLM on LDL_status
```{r}
glm_LDL <- glm(LDL_status ~  WG + RG + fibre + BMI + age + sleep_time + systolic_blood_pressure + exercise_duration, family = "binomial", data = model_data)
```

### Table 1. Logistic model with LDL-cholesterol status as the dependent variable
```{r,results="asis", eval=TRUE}
coef(summary(glm_LDL)) %>%
  as_tibble(rownames = "Variable") %>%
  rename("p Value" = "Pr(>|z|)",
         "Z Value" = "z value") %>%
  kable(digits = c(0, 3, 3, 2, 2), format = "pipe", caption = "GLM on LDL_status")
```


## Ordinal on LDL_band
```{r}
ordinal_LDL <- polr(LDL_band ~ WG + RG + fibre + BMI + age + sleep_time + systolic_blood_pressure + exercise_duration, data = model_data)
```

### Table 2. Ordinal model on LDL
```{r,results="asis", eval=TRUE}
ctable <- coef(summary(ordinal_LDL))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable <- cbind(ctable, "p Value" = p)

ctable %>% as_tibble(rownames = "Variable") %>%
  rename("t Value" = "t value",
         "Estimate" = "Value") %>%
  kable(digits = c(0, 3, 3, 2, 2), format = "pipe", caption = "Ordinal on LDL")
```

## Multinomial on LDL_band
```{r}
multinom_LDL <- multinom(LDL_band ~ WG + RG + fibre + BMI + age + sleep_time + systolic_blood_pressure + exercise_duration, data = model_data, trace=FALSE)
```

### Table 3. Multinomial model on LDL
```{r}
coef(summary(multinom_LDL)) %>% 
  as_tibble(rownames = "LDL_band") %>% 
  kable(digits = 3, format="pipe", caption = "Estimates")

# calculate p-values
zvalues <- summary(multinom_LDL)$coefficients / summary(multinom_LDL)$standard.errors
pvalues <- pnorm(abs(zvalues), lower.tail=FALSE)*2

pvalues %>%  
  as_tibble(rownames = "LDL_band") %>% 
  kable(digits = 2, format="pipe", caption = "p Values")

```

# HDL
## GLM on HDL_status
```{r}
glm_HDL <- glm(HDL_status ~  WG + RG + fibre + BMI + age + sleep_time + systolic_blood_pressure + exercise_duration, family = "binomial", data = model_data)
```

### Table 4. GLM model on HDL
```{r,results="asis", eval=TRUE}
coef(summary(glm_HDL)) %>%
  as_tibble(rownames = "Variable") %>%
  rename("p Value" = "Pr(>|z|)",
         "Z Value" = "z value") %>%
  kable(digits = c(0, 3, 3, 2, 2), format = "pipe", caption = "GLM on HDL_status")
```

## Ordinal on HDL_band
```{r}
ordinal_HDL <- polr(HDL_band ~ WG + RG + fibre + BMI + age + sleep_time + systolic_blood_pressure + exercise_duration, data = model_data)
```

### Table 5. Ordinal model on HDL
```{r,results="asis", eval=TRUE}
ctable <- coef(summary(ordinal_HDL))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable <- cbind(ctable, "p Value" = p)

ctable %>% as_tibble(rownames = "Variable") %>%
  rename("t Value" = "t value",
         "Estimate" = "Value") %>%
  kable(digits = c(0, 3, 3, 2, 2), format = "pipe", caption = "Ordinal on HDL")
```

## Multinomial on HDL_band
```{r}
multinom_HDL <- multinom(HDL_band ~ WG + RG + fibre + BMI + age + sleep_time + systolic_blood_pressure + exercise_duration, data = model_data, trace=FALSE)
```

### Table 6. Multinomial model on HDL
```{r}
coef(summary(multinom_HDL)) %>% 
  as_tibble(rownames = "HDL_band") %>% 
  kable(digits = 3, format="pipe", caption = "Estimates")

# calculate p-values
zvalues <- summary(multinom_HDL)$coefficients / summary(multinom_HDL)$standard.errors
pvalues <- pnorm(abs(zvalues), lower.tail=FALSE)*2

pvalues %>%  
  as_tibble(rownames = "HDL_band") %>% 
  kable(digits = 2, format="pipe", caption = "p Values")

```


